\input{preamble}
\begin{document}

\title{Bridgeland Stability: the General Theory}
\maketitle

Minicourse by Cristian Martinez (Universidad de los Andes, Bogotá, Colombia),
 CIMPA school Florianópolis 2025.

Notes at 
\href{http://github.com/danimalabares/cimpa-floripa}
{github.com/danimalabares/cimpa-floripa}

\bigskip\noindent

{\bf Abstract.} Bridgeland stability is a powerful tool for extracting geometry
from homological algebra. In particular, it gives a framework for studying
moduli spaces of objects in a triangulated category, such as the derived
category of an algebraic variety. The subject was born as a mathematical
interpretation of work in string theory, but has since impacted many areas,
including classical algebraic geometry, derived categories of coherent sheaves,
enumerative geometry, homological mirror symmetry, and symplectic geometry. The
goal of this course is to develop the foundations of Bridgeland stability,
covering the following topics: 1) The theory of t-structures on triangulated
categories, including examples via tilting. 2) The defniition of stability
conditions and the stability manifold, as well as Bridgeland’s deformation
theorem. 3) Constructions of stability conditions. 4) Moduli spaces of stable
objects. 

\bigskip\noindent
\tableofcontents
\bigskip\noindent

\noindent
{\bf Plan.}
\begin{enumerate}
\item Lecture 1: What is a stability condition.
\item Lecture 2: Main constructions in dimensions 2 and 3.
\item Lecture 3: Reaching the Geiseker chamber.
\item Lecture 4: Some applications on structures.
\end{enumerate}

\section{Recall}
\label{section-recall}

Recall that for $\mu_H$-stability we have the following property.

\begin{itemize}
\item (Schur's lemma.) If $A,B \in \text{Coh}$ are $\mu_H$ semistable with
$\mu_H(A)>\mu_H(B)$ then $\Hom(A,B)=0$. Exercise.
\item (Harder-Narasimhan filtrations.) $E \in \text{Coh}(X)$ then is a
filtration
$$
\xymatrix{
0\ar[r]&E_0\ar[rr]^{i_0}&&E_1\ar[dl]\ar[r]&\cdots\ar[r]
&E_{k-1}\ar[rr]^{i_k}&&E_k=E\ar[dl]\\
&&F_1& & & & F_k
}
$$
\iffalse$$
\xymatrix{
0\ar[r]&  E_0 \ar@{^{(}->}[r]&E_1\ar[d]\ar[r]&  \cdots \ar@{^{(}->}[r]&
E_k\ar[d]=E\\
& & F_1 & & F_k
}
$$\fi
with $E_0$ is the torsion subsheaf of $ E$, $F_j$ are $\mu_H$-semistable with
$$
\mu_H(F_1)>\mu_H(F_2)>\ldots>\mu_H(F_k)
$$
\end{itemize}

Exercise: (1) $\implies$ (2).

These two results basically say that $\text{Coh}(X)$ is generated by
$\mu_H$-semistable sheaves (via extensions) and torsion sheaves.
$$
0\to E_{j-1}\hookrightarrow E_j\to F_j\to_0
$$
$E_0:=F_0$.
 
Can we do this on $D^b(X)$?

We look for a subcategory with a fixed slope to generate the whole category…

\medskip

The following definition is intended to ``impose'' Schur's lemma on our objects:

\begin{definition}
\label{definition-slicing}

A {\it slicing} $\mathcal{P}$ of $D^b(X)$ consists of full subcategories
$\mathcal{P}(\phi)$ for every $\phi \in \mathbb{R}$ such that
\begin{enumerate}
\item $\mathcal{P}(\phi)[1]=\mathcal{P}(\phi+1)$ /
\item If $\phi_1>\phi_2$ then $\Hom(\mathcal{P}(\phi_1),\mathcal{P}(\phi_2))=0$.
\item (HN property.) For every $E \in D^b(X)$ there is a diagram (filtration)
$$
\phi_1>\phi_2>\cdots>\phi_k
$$
$$
\xymatrix{
0=E_0\ar[rr]^{i_0}&&E_1\ar[dl]\ar[r]&\cdots\ar[r]
&E_{k-1}\ar[rr]^{i_k}&&E_k=E\ar[dl]\\
&F_1& & & & F_k
}
$$
where $F_j=\text{Cone}(i_j) \in \mathcal{P}(d_j)$. {\bf Exercise:} this is
unique because of (2).

\item  $F \in Q(\phi)$ is called {\it semistable} of phase $\phi$. The simple objects
(those which have no subobjects) in $\mathcal{P}(\phi)$ are called {\it stable
of phase} $\phi$.

\item  $\phi_1:=\phi^+(E)$, $\phi_k:=\phi^-(E)$.
\end{enumerate}
\end{definition}

The following definitions is related to the former, and to Schur's lemma:

\begin{definition}
\label{definition-heart-of-bounded-t-structure}
The {\it heart of a bounded $t$-structure} on $D^b(X)$ is a full additive
subcategory $\mathcal{A} \subset D^b(X)$ such that
\begin{enumerate}
\item For $i>j$, $\Hom(A[i],B[j])=0$ for $A,B \in \mathcal{A}$.
\item For every $E \in D^b(X)$ there are integers $k_1>k_2>\ldots>k_m$ and a
filtration
\begin{equation}
\label{equation-diagram-abelian-category}
\xymatrix{
0=E_0\ar[rr]^{i_0}&&E_1\ar[dl]\ar[r]&\cdots\ar[r]
&E_{k-1}\ar[rr]^{i_k}&&E_k=E\ar[dl]\\
&A_1[k_1]& & & & A_k[k_m]
}
\end{equation}
exact triangles, with $A_j \in \mathcal{A}$.
\end{enumerate}
\end{definition}

\begin{exercise}
\label{exercise-A-is-abelian}
$\mathcal{A}$ is abelian.
\end{exercise}

\begin{exercise}
\label{exercise-}
If $\mathcal{P}$ is a slicing of $D^b(X)$ then $\mathcal{P}(0,]$ (the extension
closure of $\{\mathcal{P}(\phi):0<\phi \leq 1\}$) is the heart of a bounded
$t$-structure.
\end{exercise}

\begin{example}
\label{example-coherent-sheaves-are-the-heart-of-a-bounded-structure}
$\text{Coh}(X)\subset D^b(X)$ is the heart of a bounded structure.

$$
\xymatrix{
\sigma_{\leq i}(E^\bullet):\ar[d]& \cdots\ar[r]& E^{j-2}\ar[d]_=\ar[r]& 
E^{j-1}\ar[d]_=\ar[r]^{\delta}& \Ker \delta_j\ar[d]\ar[r]&  0 \ar[d]\ar[r]
&  \cdots\\
\sigma_{\leq j+1}(E^\bullet):\ar[d]&  \cdots \ar[r]& E^{j-2}\ar[d]\ar[r]
& E^{j-1}\ar[d]\ar[r]^{d_j}&E^j\ar[d]\ar[r]& \Ker(\delta_{j+1})\ar[d]\\
&& 0& 0&E_2k/\Ker \delta_j& \Ker \delta_{j+1}
}
$$
\medskip
$$
\xymatrix{
\cdots\ar[r]&\sigma_{\leq j}(E^\bullet)\ar[rr]& &\sigma_{\leq
j+1}(E^\bullet)\ar[dl]\ar[r]&  \cdots\\
&  &  H^{i+j}(E^\bullet)[-j-1]
}
$$
\end{example}

\begin{definition}
\label{definition-cohomologies}
The objects $A_j$ appearing in Eq. \ref{equation-diagram-abelian-category} are
called the {\it cohomologies} of $E$ with respect to the $t$-structure
$\mathcal{A}$, $A_j:=\mathcal{H}_{\mathcal{A}}^{-k_j}(E)$.
\end{definition}

\section{Constructing new hearts}
\label{section-constructing-new-heats}

How can we construct other $t$-structures? Suppose that $\mathcal{A}$ is an
abelian category. A {\it torsion pair} on $\mathcal{A}$ consists of two full
subcategories $(\mathcal{T}, \mathcal{F})$ such that
\begin{enumerate}
\item $\Hom(T,F)=0$ for all $T \in \mathcal{T}$ and $F \in \mathcal{F}$.
\item For every $A \in \mathcal{A}$ there is an exact sequence
$$
\xymatrix{
0\ar[r]&T\ar[r]&A\ar[r]&F\ar[r]&0
}
$$
with $T \in \mathcal{T}$ and $F \in \mathcal{F}$.
\end{enumerate}

Again: with the first property we may prove that the exact sequence in the
second property is unique (Exercise).

\begin{exercise}
\label{exercise-torsion-pair}
$\mathcal{A}=\text{Coh}(X)$, $\mathcal{T}=$ torsion sheaves, $\mathcal{F}=$
torsion free sheaves.
\end{exercise}

\begin{proposition}
\label{proposition-}
If $\mathcal{A}$ is a hear of a bounded $t$-structure on $D^b(X)$ and
$(\mathcal{T},\mathcal{F})$ is a torsion pair on $\mathcal{A}$, then
\begin{align*}
\mathcal{A}^\sharp&=\left<\mathcal{F}[1],\mathcal{T}\right>\\
&=\left\{E\in D^b(X):\substack{\mathcal{H}_{\mathcal{A}}^j(E)=0\\
\text{for $j \neq -1,0$} \\
\mathcal{H}_{\mathcal{A}}^{-1}(E)=\mathcal{F}\\
\mathcal{H}_{\mathcal{A}}^\bullet(E) in \mathcal{T} }\right\}
\end{align*}
which is called the {\it tilted heart category}.
\end{proposition}

\noindent
{\bf Note.} We can think of those as
complexes $E ^\bullet$ that fit in an exact triangle
 $$
\xymatrix{
0\ar[r]&F[1]\ar[r]&E^\bullet\ar[r]&T\ar[r]&0
}
$$
(which a prori is just an exact triangle, but it is an exact
sequence in our case).

The first proof of this fact was done by Bridgeland [confirm]:

\begin{proof}[Idea of proof]
[Picture.]
\end{proof}

We show an example to understand how the proof works.

\begin{example}
\label{example-}
Suppose that the $\mathcal{A}$-filtration of an object $E$ is of the form
$$
\xymatrix{
0\ar[r]&F_1[1]\ar[r]&E\ar[r]&F_2\ar[r]&0
}
$$

Let's comute the cohomologies:
$$
\xymatrix{
0=E_0\ar[rr]^{i_0}&&E_1\ar[dl]\ar[rr]^{i_2}&&E_2=E\ar[dl]\\
&F_1[1]&  & F_2
}
$$

$F_1,F_2 \in \mathcal{A}$. So, the cohomologies, by definition are
 $\mathcal{H}_{\mathcal{A}}^{-1}(E)=F_1$,
$\mathcal{H}_{\mathcal{A}}^0(E)=F_2$.
 
Now consider the following diagram:
$$
\xymatrix{
&  F[1]\ar[d]&  &  T\ar[d]\\
0 \ar[r]&  F_1[1]\ar[d]\ar[r]&  E \ar[r]\ar[rd]_f&  F_2\ar[d]\ar[r]& 0\\
& F[1]& & F
}
$$
so wer consider
$$
\xymatrix{
& & F[1]\ar[dl]\ar[d]\\
0\ar[r]&C=\text{Cone}(f)[-1]\ar[r]&E\ar[r]^f&F\ar[r]&0
}
$$
\end{example}

Recall that in a triangulated category,
$$
\xymatrix{
X\ar[d]\ar[r]&  0 \ar[d]\ar[r]& X[1]\ar@{.>}[d]\ar[r]& X[1]\ar[d]\\
Y\ar[r]& Z\ar[r]& X^1\ar[r]&Y[1]
}
$$
where the leftmost square commutes.

$$
\xymatrix{
0\ar[rr]&&F[1]\ar[dl]\ar[rr]^{g}&  &  C \ar[dl]\ar[rr]&  &  E\ar[dl]\\
&  \underbrace{F[1]}_{\mathcal{A}^\sharp[1]}&  &  \underbrace{\text{Cone}(g)}_{\mathcal{A}^\sharp}& &
\underbrace{F}_{\mathcal{A}^\sharp[-1]}
}
$$

\section{Stability conditions}
\label{section-stability-conditions}

\begin{definition}[Stability condition, definition 1]
\label{definition-stability-condition-1}
A {\it stability condition} is a pair $\sigma=(\mathcal{P},Z)$ where
$\mathcal{P}$ is a slicing of $D^b(X)$, and a morphism $Z$ called the  {\it
central change} of the form
$$
\xymatrix{
Z:K_0(D^b(X))\ar[rr]\ar@{->>}[dr]_\nu&  &  \mathbb{C}\\
&  \Gamma \ar[ur]_z
}
$$
where $\Gamma$ is a finite rank lattice, satisfying that
\begin{enumerate}
\item $Z(E) \in \mathcal{E}_{>0}e^{i\pi \phi}$ for any nonzero object $E \in
\mathcal{P}(\phi)$.
\item (Support property.) 
For a fixed norm $\|\cdot\|$ on $\Gamma \otimes \mathbb{R}$ we have
$$
C_{\sigma}=\text{inf}\left\{ \frac{|Z(E)|}{\|\nu(E)\|}:\substack{
\text{$E$ non-zero} \\ E \in \mathcal{P}(\phi),\phi \in \mathbb{R}} \right\} >0
$$
\end{enumerate}
\end{definition}

\begin{definition}[Stability condition, definition 2]
\label{definition-stability-condition-2}
A {\it stability condition} is a pair $\sigma=(\mathcal{A},Z)$, where
$\mathcal{A}$ is the heart of a bounded $t$-structure on $D^b(X)$, 
and a morphism $Z$ of the form
$$
\xymatrix{
Z:K_0(D^b(X))\ar[rr]\ar@{->>}[dr]_\nu&  &  \mathbb{C}\\
&  \Gamma \ar[ur]_Z
}
$$
where $\Gamma$ is a finite rank lattice, satisfying that
\begin{enumerate}
\item $Z(E) \in \mathbb{R}_{>0}e^{i\pi \phi}$ for $E \in \mathcal{A}$,
$$
\left[ \frac{\text{Im}Z(E)=0}{\text{Im}Z(E)\geq 0}\implies 
\text{Re}Z(E)<0 \right] 
$$
\item If $\mu_Z=-\frac{\text{Re}Z}{\text{Im}Z}$ then objects in $\mathcal{A}$
have the HN property w.r.t. $\mu_Z$ and we have the support property.
\end{enumerate}
\end{definition}

\begin{exercise}
\label{exercise-definitions-are-equivalent}
Prove that the two definitions of stability are equivalent.
\begin{itemize}
\item Prove Schur's Lemma for $\mu_Z$-stability.
\item $\mathcal{P}(\phi)=\mu_Z$-semistable objects of phase $\phi$, i.e., $Z(E)
\in \mathbb{R}_{>0}e^{i\pi\phi}$ for $\phi \in (0,1]$ and set
$\mathcal{P}(\phi+1)=\mathcal{P}(\phi)[1]$.
\item Take $\mathcal{A}(0,1]=$generated by $\mathcal{P}(\phi)$ with 
$0<\phi\leq 1$.
\end{itemize}
\end{exercise}

\begin{exercise}
\label{exercise-coherent-sheaves-stability-condition}
If $\dim X=1$, $\mathcal{A}=\text{Coh}(X)$, $Z= - \text{deg}+i \text{rank}$, 
$v=(\text{deg},\text{rank})$.
\end{exercise}

\medskip\noindent
{\bf Rephrasing of support property.} There exists a quadratic form $Q$ on
$\Gamma \otimes \mathbb{R}$ such that $Q(v(E))\geq 0$ for $E$
$\mu_Z$-semistable, If $Z(V)=0$ then $Q(V)<0$ for nonzero $V$.
This is like imposing Bogomolov inequality. (See moduli-spaces-of-sheaves.)

\medskip\noindent
Let $\text{Stab}^\Gamma(X)$ be the set of stability conditions with central
charge factoring through $v:K_0(\mathcal{A})\to\Gamma$.

\begin{theorem}[Bridgeland]
\label{theorem-Bridgeland}
$\text{Stab}^\Gamma(X)$ is a complex manifold. Moreover,
\begin{align*}
\text{Stab}^\Gamma(X) &\longrightarrow \Hom(\Gamma,\mathbb{C}) \\
\sigma=(\mathcal{A},Z) &\longmapsto Z
\end{align*}
is a local homomorphism.
\end{theorem}

This is known as Bridgeland deformation result. In fact, it is a consequence of
the existence of the quadratic form [reference?].

For every object in $\mathcal{A}$ where $\sigma=(\mathcal{A},Z)$ is a stability
condition there is a HN filtration
$$
\xymatrix{
0\ar[r]&E_0\ar[rr]^{i_0}&&E_1\ar[dl]\ar[r]&\cdots\ar[r]
&E_{m-1}\ar[rr]^{i_m}&&E_m=E\ar[dl]\\
&&F_1& & & & F_m
}
$$
$F_1 \in \mathcal{P}(\phi_1)$, $\phi_1:=\phi^+_\sigma(E)$, $F_m \in
\mathcal{P}(\phi_m)$, $\phi_m:=\phi_\sigma^-(E)$.

The topology on $\text{Stab}^\Gamma(X)$ is the coarsest topology that makes the
functions
$$
\sigma \mapsto \phi^+_\sigma,\qquad \sigma \mapsto \phi^-_\sigma, \qquad 
\sigma \mapsto Z
$$
[continuous].

\medskip\noindent
Now let $\dim X=2$. Let $H,B \in \text{NS}(X)_\mathbb{R}$, $H$ ample.
$$
\text{ch}^B(E)=e^{-B}\text{ch}(E)=\text{``$\text{ch}(E\otimes (-B))$''}
$$
When computing the Chern characters we obtain
\begin{align*}
\text{ch}_0^B = \text{ch}_0,&\qquad \text{ch}_1^B=\text{ch}_1-B\text{ch}_0\\
\text{ch}_2^B&=\text{ch}_2-B\text{ch}_1+\frac{B^2}{2}\text{ch}_0 
\end{align*}

\medskip\noindent
{\bf $B$-Twisted Mumford slope.}
$$
\mu_{B,H}=
\begin{cases}
\frac{H \text{ch}_1^B}{H^2 \text{ch}_0}\qquad &\text{if }\text{ch}_0\neq 0 \\
+\infty\qquad &\text{if }\text{ch}_0=0
\end{cases}
$$
\begin{align*}
\mu_{B,H}(E)&=\frac{H(\text{ch}_1(E)-B\text{ch}_0(E)}{H^2\text{ch}_0(E)}\\
&=\mu_H(E)-BH
\end{align*}
$$
\mathcal{T}_{B,H}=\{E \in \text{Coh}(X):\mu_H\text{-HN factors have }
\mu_{B,H}>0\}
$$
$$
\mathcal{F}_{B,H}=\{E \in \text{Coh}(X): \mu_H\text{-HN factors have
}\mu_{B,H}\leq 0\}
$$
\begin{itemize}
\item $\Hom(T,F)=0$ for $T \in \mathcal{T}_{B,H}$, $F \in \mathcal{F}_{B,H}$ 
(Schur's lemma).
\item If $\mathcal{E}\in\text{Coh}(X)$, take its HN-filtration
$$
\xymatrix{
0\ar[r]&E_0\ar[rr]^{i_0}&&E_1\ar[dl]\ar[r]&\cdots\ar[r]
&E_{m-1}\ar[rr]^{i_m}&&E_m=\mathcal{E}\ar[dl]\\
&&F_1& & & & F_m
}
$$
$$
\mu_{B,H}(E_0)>\mu_{B,H}(F_1)>\ldots>\mu_{B,H}(F_m)
$$
Let $j$ such that $\mu_{B,H}(F_j)>0$ and $\mu_{B,H}\leq 0$.
$$
0 \to \underbrace{E_j}_{\mathcal{T}_{B,H}}\to \mathcal{E} 
\to \underbrace{\mathcal{E}/E_j}_{\in \mathcal{F}_{B,H}}\to 0
$$
\end{itemize}

Set a tilted heart $\text{Coh}^{B,H}(X)=\left<\mathcal{F}_{B,H}[1],
\mathcal{T}_{B,H}\right>$.

Generators:
\begin{itemize}
\item $T$ torsion sheaves $\to H\text{ch}_1^B(T)\geq 0$.
\item $\mathcal{E}$ $\mu_{B,H}$-semistable sheaves with $µ_{B,H}>0 \to 
H\text{ch}_1^B (\mathcal{E})>0$.
\item $F[1]$,  $F$ is $\mu_{B,H}$-semistable with $µ_{B,H}\leq 0$.
\end{itemize}
Then
$$
H\text{ch}_1^B (F)\leq 0,\qquad H\text{ch}_1^B (F[1])\geq 0
$$
since shifting by 1 changes the sign of the Chern character. [Comment about
imaginary part of stability condition.]

Is there a central charge $Z$ such that $\sigma_{B,H}=(\text{Coh}^{B.H}(X),Z)$
is a stability condition?
$$
Z_{B,H}=\text{Re}Z_{B,H}+i H\text{ch}_1^B
$$
We need: $H\text{ch}_1^B(E)=0$ for some $E \in \text{Coh}^{B,H}(X)$ then
$\text{Re}Z_{B,H}(E)<0$.

Ingredients:
\begin{enumerate}
\item (Hodge index theorem.) $(DH)^2 \geq D^2 H^2$.
\item (Bogomolov inequality.) If $E$ is $\mu_{B,H}$-semistable then
\begin{align*}
\Delta(E)&=[\text{ch}_1(E)]^2-2\text{ch}_0(E)-\text{ch}_2(E)\geq 0\\
&=\text{ch}_1^B(E)^2-2\text{ch}_0^B\text{ch}_2^B(E)\geq 0
\end{align*}
We obtain the {\it BH-discriminant} 
$$
\Delta_{B,H}(E):=(H\text{ch}_1^B(E))^2-2\text{ch}_0(E)H^2\text{ch}_2^B(C)
\geq H^2\Delta(E)\geq 0
$$
If $E$ is $\mu_H$-semistable and $\text{Im}Z_{B,H}(E)=0$, $H\text{ch}_1^B(E)=0
\implies \text{ch}_2^B(E)\leq 0$, $\text{ch}_2^B(E)-a\text{ch}_0(E)<0$ 
for any $a>0$.
\end{enumerate}

\begin{theorem}[Bridgeland, Areni-Bestum]
\label{theorem-Bridgeland-Areni-Bersum}
 $$
Z_{B,H}=-(\text{ch}_2^B-a\text{ch}_0)+iH\text{ch}_1^B
$$
is the central charge of a stability condition on $\text{Coh}^{B,H}(X)$ 
for any $a>0$.
\end{theorem}

\begin{theorem}
\label{theorem-BH-discriminant}
$\Delta_{B,H}\geq 0$ for any $v_{B,H,a}$-semistable object, where 
$v_{B,H,a}=\frac{\text{ch}_2^B-a\text{ch}_0}{H\text{ch}_1^B}$.
\end{theorem}

\medskip\noindent
For $n:=\dim X\geq 2$,
$$
\mu_{B,H}=
\begin{cases}
\frac{H^{n-1}\text{ch}_1^B}{\text{ch}_0}\qquad &\text{if }\text{ch}_0 \neq 0 \\
+\infty\qquad &\text{if }\text{ch}_0=0
\end{cases}
$$
$\text{Coh}^{B,H}(X)$, $\Delta_{B,H}(E)=\dim n$ analogy. $Z_{B,H,a}$ does not
satisfy the positivity property of a central charge since
$$
Z_{B,H,a}=-(H^{n-2}\text{ch}_2^B-a\text{ch}_0+iH^{n-1}\text{ch}_1^B)
$$
[skyscraper sheaf argument]

\begin{exercise}
\label{exercise-E-is-a-sheaf}
Let $E \in \text{Coh}^{B,H}(X)$.
If $\text{ch}_0(E)>0$ and $E$ is $\nu_{B,H,a}$-semustable for all $a\gg 0$, then
$E$ is a sheaf. Moreover, $E$ is $\mu_{B,H}$-semistable.
\end{exercise}

\begin{proposition}[Lo,---]
\label{proposition-}
If $a>\mu_{B,H}(E)\Delta_{B,H}(E)$, then if $E$ is {\it twisted} semistable
sheaf then $E$ is $\nu_{B,H,a}$-semistable.
\end{proposition}

\medskip\noindent
If $\dim X=3$ consider the subcategories of $\text{Coh}^{B,H}(X)$
$$
T_{B,H,a}=\{E \in \text{Coh}^{B,H}(X):\text{HN factors with }
\nu_{B,H,a}>0\}
$$
$$
\mathcal{F}_{B,H,a}=\{E \in \text{Coh}^{B,H}(X):\text{HN factors with }
\nu_{B,H,a}\leq 0\}

$$
$$
\mathcal{A}=\left<\mathcal{F}_{B,H,a}[1],\mathcal{T}_{B,H,a}\right>
$$
supports a stability condition.
\end{document}
